---
title: "MODELLING GDPR VIOLATIONS WITH TIDY MODELS"
author: "christopher okoth"
date: "4/30/2020"
output:
  pdf_document: default
  html_document: default
editor_options: 
  chunk_output_type: console
---
## EXPLORE THE DATA 
```{r}
# gdpr_violations <- readr::read_tsv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-04-21/gdpr_violations.tsv')
gdpr_violations <-read.csv("gdpr_fines.csv")
library(tidyverse)
library(magrittr)
library(lubridate)
# gdpr_violations %<>% rename(country=name)
# gdpr_violations %<>% mutate(date=mdy(date))
# gdpr_violations %<>%mutate(date=na_if(date,"1970-01-01"))#possibly will leave these ones 
```

### some notes 
**Article 5:** principles for processing personal data (legitimate purpose )
**Article 6:** lawful processing of personal data ie consent etc 
**Article 13** inform subject if personal data is collected 
**Article 15:** right of access of data by subject 
**Article 32:** security of data processing (breach) - you have to process people's data securely 
```{r}
gdpr_violations %>% count(article_violated,sort = T) %>% top_n(10) %>% knitr::kable(align = "c")#the most violated articles 
gdpr_violations %>% separate_rows(article_violated,sep = "\\|") %>% count(article_violated,sort = T)#the actual number 
```

```{r}
gdpr_violations %>% ggplot(aes(price+1))+geom_histogram()+scale_x_log10()
```

```{r}
gdpr_cleaned <- gdpr_violations %>% transmute(id,country,price,
                              article_violated,
                              articles=str_extract_all(article_violated,pattern = "Art. \\d+|Art.\\d+")) %>% # use the | to get those without the pattern 
  mutate(total_articles =map_int(articles,length)) %>% #parse the column to map and return an integer length of that list 
  unnest(articles) %>%add_count(articles) %>% filter(n>10) %>% 
  select(-n)#basically just remove that new column 
#the data is now not in one violation per row but in article per row 
```

```{r}
library(ggbeeswarm)
gdpr_cleaned %>% ggplot(aes(articles,price))+
  geom_beeswarm(priority = "random")+scale_y_log10(labels=scales::dollar_format())

gdpr_cleaned %>%mutate(articles=str_replace_all(articles,pattern = "Art. ","Article") ,
                       articles=fct_reorder(articles,price)) %>% #the default function that the reoder uses is median 
  ggplot(aes(articles,price,color=articles))+geom_boxplot(alpha=0.2)+
  geom_quasirandom()+scale_y_log10(labels=scales::dollar_format())+ggdark::dark_theme_gray()+theme(legend.position="None")+
  ggtitle("Article 32 seems to have the largest fines associated with it ")

```

```{r}
gdpr_cleaned %>% mutate(value=1) %>% select(-article_violated) %>% 
  pivot_wider(names_from = articles,values_from = value,values_fn = list(value = min),values_fill = list(value=0)) %>% janitor::clean_names()->gdpr_articles
```


### do we have evidence that violating multiple articles is associated with higher fines 

## BUILD THE MODEL 

```{r}
library(tidymodels)
gdpr_recipe <- recipe(price~.,data=gdpr_articles) %>% 
  step_other(country) %>% 
  update_role(id,new_role = "id") %>% 
  step_dummy(all_nominal())
```


```{r}
gdpr_prep <- prep(gdpr_recipe)
juice(gdpr_prep)
```

```{r}
gdpr_workflow <- workflow() %>% add_recipe(gdpr_recipe) %>% 
  add_model(linear_reg() %>% set_engine("lm"))
```


## EXPLORE THE RESULTS 
```{r}
#then we use the fit dunction to fit the model using the wflow 
gdpr_workflow %>% fit(data=gdpr_articles)->gdpr_fit
#since the above is a workflow object we have to pull stuf out of it 
gdpr_fit %>% pull_workflow_fit() %>% tidy() %>% filter(p.value<0.5)
```
The more articles one violates the higher the fines one pays 
And those who violate article 15 get the highest fines 



### prediction on new data 

```{r}
new_data <- crossing(country="Other",
                     art_5=0:1,
                     art_15=0:1,
                     art_6=0:1,
                     art_32=0:1,
                     art_13=0:1) %>% mutate(total_articles=art_5+art_15+art_6+art_32+art_13,id=row_number())
new_data
```






