---
title: "random_forest_model_hyperparameter_tuning"
author: "christopher okoth"
date: "4/21/2020"
output: html_document
---
```{r}
# food_consumption <- read.csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-18/food_consumption.csv")
food_consumption <- read.csv("food_consumption.csv")
```


```{r fig.height=6,fig.width=9}
library(ggthemes)
library(tidyverse)
food_consumption %>% ggplot(aes(x=food_category,y=co2_emmission))+geom_boxplot(fill="navy")+
  theme_economist()+theme(axis.text.x = element_text(angle = 90,hjust = 1),plot.background = element_rect("gold3"))
```

Beef has the highest $CO-2$ consumption 

let us plot food consumption per capita 
```{r}
food_consumption %>% ggplot(aes(x=food_category,y=consumption))+geom_boxplot(fill="red")+coord_flip()+theme_economist()
#+theme(axis.title.x = element_text()+ xlab("Food_category")+ylab("consumption"))
```

we then get the country with the highest $CO_2$ emissions 


```{r fig.height=15}
food_consumption %>% ggplot(aes(x=country,y=co2_emmission))+
 geom_bar(stat = "identity",fill="navy")+coord_flip()

food_consumption %>% mutate(new_country=fct_lump(country,5))%>% ggplot(aes(x=new_country,y=co2_emmission))+
 geom_bar(stat = "identity",fill="navy")+coord_flip()
```

## Sum emmissions per country and get the top 10


```{r}
sum_emmissions <- food_consumption %>% group_by(country) %>% 
  summarise(all_emmission=sum(co2_emmission))
sum_emmissions_order <- arrange(sum_emmissions,desc(all_emmission))

```

```{r}
knitr::kable(head(sum_emmissions_order,n=10))
```


# But all that ca be accomplished in a single line of code 
```{r}
food_consumption %>% count(country,wt=co2_emmission,sort = T) %>% head(10)->top10
```


## then we plot 


```{r}
selected <- food_consumption %>% 
  filter(country%in%sum_emmissions_order$country[1:10])#getting all the information about the top10 from the parent data 
#mtcars$cyl%in%6

```
```{r}
# to order make factors 
selected$country <- factor(selected$country,levels = c("Argentina",		
"Australia",		
"Albania	",		
"New Zealand",	
"Iceland",			
"USA",			
"Uruguay",	
"Brazil",		
"Luxembourg",		
"Kazakhstan"))
```


```{r fig.height=6,fig.width=10}
selected %>% ggplot(aes(x=country,y=co2_emmission))+geom_bar(stat = "identity",fill="coral1")+theme_economist()+
  theme(axis.text.x = element_text(angle =90,hjust = 1),plot.background = element_rect("aliceblue"))+facet_wrap(.~food_category)+coord_flip()
```


```{r eval=FALSE, include=FALSE}
food_consumption %>% filter(country %in% top10$country) %>% ggplot(aes(factor(country),co2_emmission,fill=country))+geom_col()+
  facet_wrap(~food_category)+coord_flip()+ggsci::scale_fill_simpsons()
```






### then we can go ahead for a radnom forest model hyperparameter tuning 

```{r}
#but first we explore our data to see just how it looks like 
library(countrycode)
food_consumption %>% mutate(continent=countrycode(sourcevar = country,origin = "country.name",destination = "continent"))

food_consumption %>% mutate(continent=countrycode(sourcevar = country,origin = "country.name",destination = "continent")) %>% count(continent,wt=co2_emmission,sort = T)
```
Europe is by far the leading continent in co2 emmissions followed by the Americas 


```{r}
#then we can have the data in a wider format just so we know how much pork does each country consume and so on and so fort 
food_consumption %>% mutate(continent=countrycode(sourcevar = country,origin = "country.name",destination = "continent")) %>% select(-co2_emmission) %>% pivot_wider(names_from = food_category,values_from = consumption) %>% 
  janitor::clean_names()
```

### then let us  a model to predict whether a country is in Asia or not depending on its pattern of food consumption
```{r}
food_consumption %>% mutate(continent=countrycode(sourcevar = country,origin = "country.name",destination = "continent")) %>% select(-co2_emmission) %>% count(continent,wt=consumption,sort = T)
```

```{r}
#we can create a asia column 
# food_consumption %>% mutate(continent=countrycode(sourcevar = country,origin = "country.name",destination = "continent")) %>% select(-co2_emmission) %>% 
#   mutate(Asia=ifelse(continent=="Asia","Asia","Other"))


#alternatively
food_consumption %>% mutate(continent=countrycode(sourcevar = country,origin = "country.name",destination = "continent")) %>% select(-co2_emmission) %>% pivot_wider(names_from = food_category,values_from = consumption) %>% 
  janitor::clean_names()%>%mutate(Asia=case_when(continent=="Asia"~"Asia",T~"Other")) %>% 
  select(-continent,-country) %>%
  #then we change everything else that is a character to a factor
  mutate_if(is.character,factor)->food
```

So asia is what we are going to predict 

And since every column apart from asia is numeric we can do a scatter plot matrix to just have a quick glance at our data 


```{r}
library(GGally)

ggscatmat(food,columns = 1:11,color = "Asia",alpha=0.7)
```
So what does it tell us for instance in the rest of the world places with more beef havve more puoltry but in Asia that is hardly the case , and in the case of rice there is a signinficant difference in that lots of people eat rice but less so in the rest of the world 

So a random forest model will be very useful since there are different parameters 

#then let us go ahead and tune hyperparameters 
```{r}
library(tidymodels)
#then we create some bootstap samples 
set.seed(1234)
food_boot <- bootstraps(food,20)#this has been done with replacement and it is the same sixe as our original dataset and so there are duplicates in it   
```
and the assesment part of our split is that which did not make it to our bootstrap . the bootsrap are built with random samples and it has randoms sample for the different set of countries 


#then onto random forest 

```{r}
#so since we want to classify whether or not a country is in Asia or not 
model_specification <- rand_forest(mode="classification",mtry = tune(),min_n = tune(),trees = 1000) %>% 
  set_engine("ranger")#the tidy models framework 
```


when we train models  there are parameters that we can learn from the data . but there are those which we cannot , these are the hyperparameters 

in a random forest consider trees that are voting 

so mtry is how many predictors will be used randomly at each split 
min_n is the number of data points in a node that is required for that node to be split further 
These are parametrs we have to put ahead of time 

you can either used best practices or you can tune your model


```{r}
#now it is time to go to tuning 

tune_rf <- tune_grid(
  Asia~.,model = model_specification,resamples = food_boot
)
```

one can use different approaches to how you are going to search for these parameters 


This thing is training a random forest model with every combination of mtry and min_n reasonable from the data and is fittinig every combination 20 times since that is the number of bootstraps 

```{r}
tune_rf %>% collect_metrics()
```
```{r}
#then we can define the best one and specify by what metric we want this done 
tune_rf %>% show_best(metric = "roc_auc")
```
## just a side not one can use parallel processing to do the tuning in the tidymodels
